---
title: "Nyack Floodplain Data Archiving"
author: "Amalia Handler"
date: "6/24/2019"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

Code to compile all River Net Nyack Floodplain meterological and hydrologic data. Code was written by Amalia Handler, Environmental Data Initiative Summer Data Science Fellow, during summer 2019. Please contact her with any questions: amalia.handler@gmail.com

Load data from FLBS internal database. It extends from 2013-04-01 to 2019-06-25. 

```{r include = FALSE}
# List the files in the directory from Phil
ha_files_toread <- list.files(path = "/Users/Sam Bosio/Documents/FLBS_NyackFloodplain/Data/Nyack_RiverNET2020", pattern = ".dat")

# Read in the files
toread <- paste("/Users/Sam Bosio/Documents/FLBS_NyackFloodplain/Data/Nyack_RiverNET2020/", ha_files_toread, sep = '')

ha_files <- list()
ha_files <- lapply(toread, read.delim, header = FALSE, sep = ',', skip = 4) 

# Get the column names
var_names <- list()
for(i in 1:length(ha_files)){
  var_names[[i]] <- colnames(read.delim(paste("./Data/", ha_files_toread[i], sep = ''), sep = ",", skip = 1, header = TRUE))
}

# Apply the column names to the list of dfs
for(i in 1:length(var_names)){
  colnames(ha_files[[i]]) <- var_names[[i]]
}

# Convert the timestamp to a datetime variable
ha_files <- lapply(ha_files, function(x){
 x$TIMESTAMP <- as.POSIXct(x$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S")
 return(x)
})

# Determine if each file is a water or meteorlogical file
file_info <- sapply(ha_files, function(x){
  if(colnames(x)[3] == "AirTC") {
    data_type <- "met"
  } else { data_type <- "water"}
  return(data_type)
})

# Find the ha number for each file
# If the file contains "CR1000", then I can extract
# If the file name contains "Nyack" then need another method
well_info <- NULL
for(i in 1:length(ha_files_toread)){
  if(grepl("CR1000", ha_files_toread[i])){
  well_info[i] <- substr(ha_files_toread[i], 8, 11)
  } else {well_info[i] <- substr(ha_files_toread[i], 7, 10)}
}

# Name the dataframes in the list
ha_df_names <- paste(well_info, file_info, sep = '_')

names(ha_files)  <- ha_df_names
names(var_names) <- ha_df_names

# Need to coerce the mutiple columns in the HA02 df from character to number. Seems that R does not see "NAN" is equivelent to "NaN".
column_numbers <- c(11, 14:17)
for(col in column_numbers){
  ha_files$HA02_water[,col] <- 
    as.numeric(ha_files$HA02_water[,col])
}

# Remove rows with no datetime information (artifact of changing over to daylight savings time)
ha_files <- lapply(ha_files, function(df){
  df[!is.na(df$TIMESTAMP),]
})


```


Isolate the met station data, this is part of it's own data package

```{r}
# Rename "timestamp" to "datetime" column and remove the records column in the met data
colnames(ha_files$HA07_met)[1]  <- "datetime"

# Add a site column, then rearrange the columns, getting rid of the RECORD number column
ha_files$HA07_met$site <- "ha07" 
ha_files$HA07_met      <- ha_files$HA07_met[, c(21,1,3:20)]

# Save the meterological data file
saveRDS(ha_files$HA07_met, './Data/Nyack_Met_Data.rds')
write.csv(ha_files$HA07_met, './Data/Nyack_Met_Data.csv', row.names = FALSE)

```


Compile all water data into one dataframe. This is part of another data package.


```{r}
# For HA02, 08, and 15, there are two sets of sensors. Need to separate these into different dataframes.
#   HA02 has sensors 9 and 10
#   HA08 has sensors 1 and 2
#   HA15 has sensors 7 and 8
ha02_09 <- ha_files$HA02_water[,colnames(ha_files$HA02_water)[1:10]]
ha02_10 <- ha_files$HA02_water[,colnames(ha_files$HA02_water)[c(1,2,11:18)]]
ha08_01 <- ha_files$HA08_water[,colnames(ha_files$HA08_water)[1:10]]
ha08_02 <- ha_files$HA08_water[,colnames(ha_files$HA08_water)[c(1,2,11:18)]]
ha15_07 <- ha_files$HA15_water[,colnames(ha_files$HA15_water)[1:10]]
ha15_08 <- ha_files$HA15_water[,colnames(ha_files$HA15_water)[c(1,2,11:18)]]

# Recombine all data into a list to ease processing
ha_sep <- list(ha_files$CASC_water, ha02_09, ha02_10, ha_files$HA07_water,
               ha08_01, ha08_02, ha_files$HA10_water, ha_files$HA12_water,
               ha15_07, ha15_08)

# Need to add site information to each df
site <- c('CASC', 'HA02', 'HA02', 'HA07', 'HA08', 'HA08', 'HA10', 'HA12', 'HA15', 'HA15')

# Also add sensor number to each df. This is just an elaborate was of identifying when the sensor number requires a leading zero (for sensors 1-9) or not (for sensor 10).
sensor <- sapply(ha_sep, function(x){
  col_name <- colnames(x)[10]
  if(nchar(col_name) == 8){
    num <- substr(col_name, 8, 8)
    sensor_number <- paste('0', num, sep = '')
  } else {
    sensor_number <- substr(col_name, 8, 9)
  }
  return(sensor_number)
})

# Add columns to each df for the site and the sensor number
ha_sep <- lapply(seq_along(ha_sep), function(i){
  ha_sep[[i]]$site          <- site[i]
  ha_sep[[i]]$sensor_number <- sensor[i]
  return(ha_sep[[i]])
})

# Now bind together all the files to form one very long dataframe.

# Need to rename columns for compatabiility
col_names <- c('datetime', 'record', 'do_conc', 'do_sat', 'do_temp', 'cond', 'ct', 'cond_temp', 'level_m', 'level_temp', 'site', 'sensor_number')

# Rename the columns for the dfs in the list
ha_rename <- lapply(ha_sep, function(x){
  colnames(x) <- col_names
  return(x)
})

# Bind together all rows to create one giant df
ha_all <- do.call("rbind", ha_rename)

# Rearrange the columns
ha_water <- ha_all[, c(11,12,1,3:10)]

# Remove duplicated rows
ha_water <- ha_water[!duplicated(ha_water),]

# Save the new master file
saveRDS(ha_water, './Data/HA_Water_Data_Compiled.rds')
write.csv(ha_water, './Data/HA_Water_Data_Compiled.csv', row.names = FALSE, quote = FALSE)

```

Voila! All River Net data are compiled.
